---
title: "Relative Resilience Calculations"
author: "Socheata Hour"
date: "January 15, 2025"
output:
  html_document: default
  md_document: default
  pdf_document: default   
  citation_package: biblatex
bibliography: relative_resilience_reference_code.bib
---

## References 
@eichristenson_eichristensonstacy-ceballos-index_2023

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
renv::init()

## Load necessary packages
```{r}
install.packages("readxl",repos = "http://cran.us.r-project.org")
install.packages("DescTools",repos = "http://cran.us.r-project.org")
install.packages("ggplot2",repos = "http://cran.us.r-project.org")
```

## Load the library package 
```{r}
library("readxl") 
library("DescTools")
library("ggplot2")
```


## How to use this code? 
+ replace your file name 
+ replace your sheet name
+ replace your limits of interest (hour post-infection)
+ run the code 

## Read in your data to the R environment from excel using Import Dataset function and perform RR calculation 
`the imported excel sheet and the result table will be display after you run the code`

Data should be in the following format in excel:

hpi=Hours post infection, abs=absorbance at 600nm

| time.hpi | ctrl.abs | virus1.abs | virus2.abs |
|----------|----------|------------|------------|
| 0        | 0.2      | 0.2        | 0.2        | 
| 2        | 0.4      | 0.3        | 0.35       |
```{r}
current_directory <- getwd()         # Find your file path at the R console by getwd()
current_directory
file_name = "example_data.xlsx"      # Replace your file name 
file_path <- file.path(current_directory, file_name)
sheet_name <- "Sheet1"               # Replace with your sheet name 

# Read in your data to the R environment from excel 
data <- read_excel(file_path, sheet = sheet_name)

# Replace the time.hpi limits
# Default: the entire growth curve
limited_row <- 1:nrow(data)          # default hpi (e.g., from 0hpi to 192hpi)

custom_limits <- TRUE                # *** TRUE for custom limits, FALSE for the default limits ***  
if (custom_limits) {
  limited_row <- 4:28                # Replace with hpi of interested (e.g., at index 4 (18hpi), index 28 (162hpi))
}

filtered_data <- data[limited_row,] 

# time.hpi difference between ith and i+1th
time_diff <- sapply(1:(nrow(filtered_data)-1), function(i) filtered_data$time.hpi[i+1]-filtered_data$time.hpi[i])
# AUC for each time.hpi 
each_auc_con <- sapply(1:(nrow(filtered_data)-1), function(i) AUC(filtered_data$time.hpi[i:(i+1)], filtered_data$ctrl.abs[i:(i + 1)]))
each_auc_1 <- sapply(1:(nrow(filtered_data)-1), function(i) AUC(filtered_data$time.hpi[i:(i+1)], filtered_data$virus1.abs[i:(i + 1)]))
each_auc_2 <- sapply(1:(nrow(filtered_data)-1), function(i) AUC(filtered_data$time.hpi[i:(i+1)], filtered_data$virus2.abs[i:(i + 1)]))

# Construct a results table and store the calculation value to this table
results <- data.frame(
  time_diff = time_diff,
  # AUC 
  auc_con = cumsum(each_auc_con),
  auc_1 = cumsum(each_auc_1),
  auc_2 = cumsum(each_auc_2),
  # Nasymptote
  Nasym_con = sapply(1:(nrow(filtered_data)-1), function(i) max(filtered_data$ctrl.abs[1:i])),
  Nasym_1 = sapply(1:(nrow(filtered_data)-1), function(i) max(filtered_data$virus1.abs[1:i])),
  Nasym_2 = sapply(1:(nrow(filtered_data)-1), function(i) max(filtered_data$virus2.abs[1:i]))
)
# PImax
results$PImax_1 <- (1-results$Nasym_1/results$Nasym_con)*100
results$PImax_2 <- (1-results$Nasym_2/results$Nasym_con)*100
# PIauc
results$PIauc_1 <- (1-results$auc_1/results$auc_con)*100
results$PIauc_2 <- (1-results$auc_2/results$auc_con)*100
# Isc
results$Isc_1 <- sqrt(abs(results$PIauc_1)*abs(results$PImax_1))
results$Isc_2 <- sqrt(abs(results$PIauc_2)*abs(results$PImax_2))
# VR
each_VR_1 <- sapply(1:(nrow(filtered_data)-1), function(i) (results$Isc_1[i+1]+results$Isc_1[i])/2*(results$time_diff[i])/(filtered_data$time.hpi[nrow(filtered_data)]-filtered_data$time.hpi[1]))
each_VR_2 <- sapply(1:(nrow(filtered_data)-1), function(i) (results$Isc_2[i+1]+results$Isc_2[i])/2*(results$time_diff[i])/(filtered_data$time.hpi[nrow(filtered_data)]-filtered_data$time.hpi[1]))
results$VR_1 <- cumsum(each_VR_1)
results$VR_2 <- cumsum(each_VR_2)
# RR
results$RR_1 <- 1/results$VR_1
results$RR_2 <- 1/results$VR_2

# Display tables in the R console 
filtered_data
results
```
## Plot the graph 
```{r}
# Find the highest y value of all the dataset 
max_y <- max(data[, !names(data) %in% "time.hpi"], na.rm = TRUE)

# Create a graph 
ggplot() + 
  geom_line(data=data, aes(x=time.hpi, y=ctrl.abs), color="black", linewidth=1) + 
  geom_line(data=data, aes(x=time.hpi, y=virus1.abs), color="darkblue", linewidth=1) + 
  geom_line(data=data, aes(x=time.hpi, y=virus2.abs), color="darkred", linewidth=1) + 
  
  # Add the limits vertical line  
  geom_vline(xintercept = data$time.hpi[head(limited_row,1)], linetype = "solid", color = "grey", alpha = 1, linewidth = .2) + # Replace the number to the limit of interested
  geom_vline(xintercept = data$time.hpi[tail(limited_row,1)], linetype = "solid", color = "grey", alpha = 1, linewidth = .2) +
  
  theme(
    legend.position = "none",
    text = element_text(family = "Times New Roman", size=12, color="black"),
    panel.background = element_blank(), 
    panel.border = element_blank(), 
    axis.line = element_line(color="black", linewidth=0.5), 
    axis.text.x = element_text(angle=90, vjust=0.5, hjust=1, color="black"),
    axis.text.y = element_text(color="black")
  ) +
  scale_x_continuous(expand=c(0, 0)) +
  scale_y_continuous(expand=expansion(mult=c(0, 0)), limits=c(0.00, max_y+0.2)) +
  labs(x="Hours Post-Infection (hpi)", y="Optical Density (OD600nm)", title="") +
  coord_cartesian(ylim=c(0.00, max_y+0.2))
```

renv::snapshot()
